### 2021-09-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@4e8ee43f38230d0f7b71ec0d6cdee36478cd10d1 🚀

\vfill
Despite having established its usefulness in the last ten years, the
decomposition of ecological networks in components allowing to measure
their \(eta\)-diversity retains some methodological ambiguities.
Notably, how to quantify the relative effect of mechanisms tied to
interaction rewiring \emph{vs.} species turnover has been interpreted
differently by different authors. In this contribution, I present
mathematical arguments and numerical experiments that should (i)
establish that the decomposition of networks as it is currently done is
indeed fit for purpose, and (ii) provide guidelines to interpret the
values of the components tied to turnover and rewiring.

@@

Ecological networks are variable both in time and space (Poisot \emph{et
al.} 2015; Trøjelsgaard \& Olesen 2016) - this variability motivated the
emergence of methodology to compare ecological networks, in a way that
meshes with the usual approaches of comparison of ecological
communities, \emph{i.e.} \(\beta\)-diversity; although the definiton of
\(\beta\)-diversity is a contentious topic amongst community ecologists
(see \emph{e.g.} Tuomisto 2010), the need to understand network
variability is motivated by the fact that species that make up the
networks do not react to their environment in the same way, and
therefore the \(\beta\)-diversity of networks may behave in complex
ways.

Poisot \emph{et al.} (2012) and Canard \emph{et al.} (2014) have
suggested an approach to \(\beta\)-diversity for ecological networks
which is based on the comparison of shared and unique links among
species, and differentiate this sharing of links between common and
unique species. This framework can be summarized as
\(\beta_{wn} = \beta_{os} + \beta_{st}\), namely the fact that overall
network dissimilarity (\(\beta_{wn}\)) has a component that can be
calculated directly from the dissimilarity of interactions between
shared species (\(\beta_{os}\)), and a component that cannot, the later
originating in unique species introducing their unique interactions
(\(\beta_{st}\)). This approach has been widely adopted since its
publication, with recent examples using it to understand the effect of
fire on pollination systems (Baronio \emph{et al.} 2021); the impact of
rewiring on spatio-temporal network dynamics (Campos-Moreno \emph{et
al.} 2021); the effects of farming on rural and urban landscapes on
species interactions (Olsson \emph{et al.} 2021); and as a tool to
estimate the sampling completeness of networks (Souza \emph{et al.}
2021). It has, similarly, received a number of extensions, including the
ability to account for interaction strength (Magrach \emph{et al.}
2017), the ability to handle probabilistic ecological networks (Poisot
\emph{et al.} 2016), and the integration into the Local Contribution to
Beta Diversity (Legendre \& De Cáceres 2013) approach to understand how
environment changes drive network dissimilarity (Poisot \emph{et al.}
2017).

In a recent contribution, Fründ (2021) argues that the calculation of
network dissimilarity terms as outlined by Poisot \emph{et al.} (2012)
is incorrect, as it can lead to over-estimating the role of interactions
between shared species in a network (``rewiring''), and therefore
underestimate the importance of species turnover across networks. Here,
I present a more thorough justification of the methodological choices
for the Poisot \emph{et al.} (2012) method, explain how information
about species turnover can be extracted from its decomposition, and
conduct numerical experiments to guide the interpretation of the
\(\beta\)-diversity values thus obtained. These numerical experiments
establish three core facts. First, the decomposition responds to the
correct sources of network variation; second, the decomposition
adequately captures the relative roles of species turnover and
interaction rewiring; finally, the decomposition adequately captures the
role of turnover vs.~non-turnover (like changes in connectance)
processes. Although the alternative normalization suggested by Fründ
(2021) is not without its uses, which I discuss in conclusion, it is
inadequate as a network \(\beta\)-diversity measurement, as it
introduces many confounding elements that make the interpretation of the
results more difficult, and should likely not be used as a default.

>>>>

\vfill
Despite having established its usefulness in the last ten years, the
decomposition of ecological networks in components allowing to measure
their \(\beta\)-diversity retains some methodological ambiguities.
Notably, how to quantify the relative effect of mechanisms tied to
interaction rewiring \emph{vs.} species turnover has been interpreted
differently by different authors. In this contribution, I present
mathematical arguments and numerical experiments that should (i)
establish that the decomposition of networks as it is currently done is
indeed fit for purpose, and (ii) provide guidelines to interpret the
values of the components tied to turnover and rewiring.

@@

Ecological networks are variable both in time and space (Poisot \emph{et
al.} 2015; Trøjelsgaard \& Olesen 2016) - this variability motivated the
emergence of methodology to compare ecological networks, including in a
way that meshes with the core concept for the comparison of ecological
communities, namely \(\beta\)-diversity (Poisot \emph{et al.} 2012). The
need to understand network variability through partitioning in
components equivalent to \(\alpha\), \(\beta\), and \(\gamma\)
diversities is motivated by the prospect to further integrate the
analysis of species interactions to the analysis of species
compositions. Because species that make up the networks do not react to
their environment in the same way, and because interactions are only
expressed in subsets of the environments in which species co-occurr, the
\(\beta\)-diversity of networks may behave in complex ways, and its
quantification is likely to be ecologically informative.

Poisot \emph{et al.} (2012) and Canard \emph{et al.} (2014) have
suggested an approach to \(\beta\)-diversity for ecological networks
which is based on the comparison of the number of shared and unique
links among species within a pair of networks. Their approach
differentiates this sharing of links between those established between
species occurring in both networks, and those established with at least
one unique species. This framework is expressed as the decomposition
\(\beta_{wn} = \beta_{os} + \beta_{st}\), namely the fact that network
dissimilarity (\(\beta_{wn}\)) has a component that can be calculated
directly from the dissimilarity of interactions between shared species
(\(\beta_{os}\)), and a component that cannot (\(\beta_{st}\)).
Presumably, the value of these components for a pair of networks can
generate insights about the mechanisms involved in dissimilarity.

This approach has been widely adopted since its publication, with recent
examples using it to understand the effect of fire on pollination
systems (Baronio \emph{et al.} 2021); the impact of rewiring on
spatio-temporal network dynamics (Campos-Moreno \emph{et al.} 2021); the
effects of farming on rural and urban landscapes on species interactions
(Olsson \emph{et al.} 2021); the impact of environment gradients on
multi-trophic metacommunities (\textbf{Ohlmann2018MapImp?}); and as a
tool to estimate the sampling completeness of networks (Souza \emph{et
al.} 2021). It has, similarly, received a number of extensions,
including the ability to account for interaction strength (Magrach
\emph{et al.} 2017), the ability to handle probabilistic ecological
networks (Poisot \emph{et al.} 2016), and the integration into the Local
Contribution to Beta Diversity (Legendre \& De Cáceres 2013) approach to
understand how environment changes drive network dissimilarity (Poisot
\emph{et al.} 2017).

This later component originates in unique species introducing their
unique interactions both between themselves, and with species that are
common to both networks. CONTINUE WITH BETA ST

Furthermore, much like the definition of \(\beta\)-diversity in all its
forms is a contentious topic amongst community ecologists (see
\emph{e.g.} Tuomisto 2010), the \(\beta\)-diversity of networks has been
submitted to methodological scrutiny over the years. A synthesis of some
criticisms, related to the correct denominator to use to express the
proportion of different links, has recently been published (Fründ 2021).
It argues that the calculation of network dissimilarity terms as
originally outlined by Poisot \emph{et al.} (2012) is incorrect, as it
can lead to over-estimating the role of interactions between shared
species in a network (``rewiring''), and therefore underestimate the
importance of species turnover across networks. As mist-understanding
either of these quantities can lead to biased inferences about the
mechanisms generating network dissimilarity, it is important to assess
how the values (notably of \(\beta_{os}\), and therefore of
\(\beta_{st}\)) react to methodological choices.

Here, I present a more thorough justification of the methodological
choices for the Poisot \emph{et al.} (2012) method, explain how
information about species turnover can be extracted from its
decomposition, and conduct numerical experiments to guide the
interpretation of the \(\beta\)-diversity values thus obtained. These
numerical experiments establish three core facts. First, the
decomposition responds to the correct sources of network variation;
second, the decomposition adequately captures the relative roles of
species turnover and interaction rewiring; finally, the decomposition
adequately captures the role of turnover vs.~non-turnover (like changes
in connectance) processes. Although the alternative normalization
suggested by Fründ (2021) is not without its uses, which I discuss in
conclusion, it is inadequate as a network \(\beta\)-diversity
measurement, as it introduces many confounding elements that make the
interpretation of the results more difficult, and should likely not be
used as a default.
